Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-08-14 paper.qmd
Replaced Six to Eight in Figure 1 title
Added axes to the projection: Need to check with the origin

```{r}
#| label: fig-NLDR-variety
#| echo: false
#| fig-cap: "Six different NLDR representations of the same data. Different techniques and different parameter choices are used. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical parameter choice. Would they make different decisions downstream in the analysis depending on which version seen? Which is the most accurate representation of the structure in high dimensions?"
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%
# (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 30), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5)
nldr1 + nldr2 + nldr3 + nldr4 +
  nldr5 + nldr6 + nldr7 + nldr8 +
  plot_layout(ncol = 4)
```

@@

>>>>

```{r}
#| label: fig-NLDR-variety
#| echo: false
#| fig-cap: "Eight different NLDR representations of the same data. Different techniques and different parameter choices are used. Researchers may have seen any of these in their analysis of this data, depending on their choice of method, or typical parameter choice. Would they make different decisions downstream in the analysis depending on which version seen? Which is the most accurate representation of the structure in high dimensions?"
#| fig-width: 8
#| fig-height: 4
#| out-width: 100%
# (a) UMAP (n_neighbors = 30, min_dist = 0.3), (b) UMAP (n_neighbors = 5, min_dist = 0.01), (c) UMAP (n_neighbors = 15, min_dist = 0.99), (f) tSNE (perplexity = 5), (g) tSNE (perplexity = 30), (l) TriMAP (n_inliers = 12, n_outliers = 4, n_random = 3), (q) PaCMAP (n_neighbors = 30, init = random, MN_ratio = 0.9, FP_ratio = 5)
nldr1 + nldr2 + nldr3 + nldr4 +
  nldr5 + nldr6 + nldr7 + nldr8 +
  plot_layout(ncol = 4)
```

@@

<!--to draw axes-->
```{r}

## Draw the axes

projection <- cbind(
  c(0.18390,-0.18443,-0.06860,0.00654,-0.02921,0.02313,-0.08963),
  c(-0.02902,0.03638,-0.16872,0.22418,0.01146,0.02777,0.01453))

projected_df <- as.matrix(true_model_df_rm_id) %*% projection

limits <- 1
rng <- range(projected_df)
projected_df <- projected_df/max(abs(rng))
colnames(projected_df) <- c("P1", "P2")

obs_labels <- as.character(1:nrow(true_model_df_rm_id))

projected_df_n <- as.matrix(true_model_df_rm_id) %*% projection |> 
  as_tibble(.name_repair = "unique") |> 
  rename(c("proj1" = "...1", 
           "proj2" = "...2"))

projected_df$obs_labels <- obs_labels

#if (position == "center") {
  axis_scale <- 2 * limits/3
  #axis_pos <- 0
  axis_pos <- -0.6 * limits
#}
# else if (position == "bottomleft") {
#   axis_scale <- limits/6
#   axis_pos <- -2/3 * limits
# }

mean_cols <- colMeans(true_model_df_rm_id, na.rm = TRUE)

adj <- function(x) axis_pos + x * axis_scale
axes <- data.frame(x1 = adj(0), 
                   y1 = adj(0), 
                   x2 = adj(projection[, 1]), 
                   y2 = adj(projection[, 2]))

# axes <- data.frame(x1 = mean_cols + mean_cols * axis_scale, 
#                    y1 = mean_cols + mean_cols * axis_scale, 
#                    x2 = mean_cols + projection[, 1] * axis_scale, 
#                    y2 = mean_cols + projection[, 2]* axis_scale)

axis_labels <- colnames(true_model_df_rm_id)
rownames(axes) <- axis_labels
theta <- seq(0, 2 * pi, length = 50)
circle <- data.frame(c1 = adj(cos(theta)), c2 = adj(sin(theta)))

ggplot() +
  geom_point(data = projected_df_n, 
             aes(
               x = proj1, 
               y = proj2),
             alpha = 0.5)  +
  geom_segment(data=axes, aes(x=x1, y=y1, xend=x2, yend=y2), colour="grey70") +
  geom_text(data=axes, aes(x=x2, y=y2, label=rownames(axes)), colour="grey50")
